---
title: "3_nutrients"
format: html
editor_options: 
  chunk_output_type: console
---
# Preliminaries

```{r}
#| label: packages
#| message: false
#| warning: false
library(tidyverse)
library(vegan)
library(mvabund)
library(knitr)
library(patchwork)
library(pairwiseAdonis) #install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
```

```{r}
#| label: data
ponchos_raw <- read.csv("data/ponchos_master.csv")
```

In order to standardize the oxygen rates, we'll apply the following conversion factors:

$$ x \frac{mmol}{m^3} =  x \frac{\mu mol}{L}\cdot \text{Volume measured} (L) \cdot  \frac{1}{time (h)} \cdot \frac{1}{\text{biomass}(g)} \cdot = x\frac{\mu mol}{g_{biomass} \cdot h}$$

In this case, since the samples were collected from the entire volume, we can use 0.550 L as the common volume for all the samples

```{r}
#| echo: false
#| message: false
nutr_names <- c("NH4", "PO4", "NO2", "NO3", "NOx", "SiO4", "DOC", "TDN", "DON")

nutr_init_avg <- ponchos_raw |> 
  filter(chamber == "environment" & is.na(O2_0)) |>
  mutate(NOx = NO2 + NO3,
         DON = TDN - NOx - NH4) |> 
  group_by(Month) |> 
  summarise(across(all_of(nutr_names), mean))

nutr_init_sd <- ponchos_raw |> 
  filter(chamber == "environment" & is.na(O2_0)) |>
  mutate(NOx = NO2 + NO3,
         DON = TDN - NOx - NH4) |> 
   group_by(Month) |> 
  summarise(across(all_of(nutr_names), sd))
```

 Because some of the data include technical replicates, let's apply a custom
 function to select the replicates based on the combinations that minimize the coefficient of variance

```{r}
#| warning: false
source("R/trepf.R")
source("R/cv_get.R")
```

Then we can proceed to calculate the flux rates

```{r}
set.seed(1)

nutr_rate <- nutr_long |> 
  filter(!(Observation %in% c("t0", "rock"))) |>
  mutate(t0 = as.POSIXct(t0, format = "%H:%M"),
         tf = as.POSIXct(tf, format = "%H:%M"),
         total_time = as.numeric(tf - t0),
         Posidonia = as.factor(Posidonia),
         Chondrilla = as.factor(Chondrilla),
         light = as.factor(light),
         dw_rhizome2 = ifelse(
           !is.na(dw_rhizome),
           dw_rhizome,
           sample(ponchos_raw[ponchos_raw$dw_rhizome>0 & !is.na(ponchos_raw$dw_rhizome),"dw_rhizome"], 24)
         ),
         delta_cr = case_when(
           ntr == "NH4" & Month == "November" ~ (conc- nutr_init_avg["NH4"][[1]][1])*(volume/(total_time)),
           ntr == "PO4" & Month == "November" ~ (conc- nutr_init_avg["PO4"][[1]][1])*(volume/(total_time)),
           ntr == "NO2" & Month == "November" ~ (conc- nutr_init_avg["NO2"][[1]][1])*(volume/(total_time)),
           ntr == "NO3" & Month == "November" ~ (conc- nutr_init_avg["NO3"][[1]][1])*(volume/(total_time)),
           ntr == "SiO4" & Month == "November" ~ (conc- nutr_init_avg["SiO4"][[1]][1])*(volume/(total_time)),
           ntr == "DOC" & Month == "November" ~ (conc- nutr_init_avg["DOC"][[1]][1])*(volume/(total_time)),
           ntr == "DON" & Month == "November" ~ (conc- nutr_init_avg["DON"][[1]][1])*(volume/(total_time)),
           ntr == "NOx" & Month == "November" ~ (conc- nutr_init_avg["NOx"][[1]][1])*(volume/(total_time)),
           ntr == "TDN" & Month == "November" ~ (conc- nutr_init_avg["TDN"][[1]][1])*(volume/(total_time)),
           ntr == "NH4" & Month == "May" ~ (conc- nutr_init_avg["NH4"][[1]][2])*(volume/(total_time)),
           ntr == "PO4" & Month == "May" ~ (conc- nutr_init_avg["PO4"][[1]][2])*(volume/(total_time)),
           ntr == "NO2" & Month == "May" ~ (conc- nutr_init_avg["NO2"][[1]][2])*(volume/(total_time)),
           ntr == "NO3" & Month == "May" ~ (conc- nutr_init_avg["NO3"][[1]][2])*(volume/(total_time)),
           ntr == "SiO4" & Month == "May" ~ (conc- nutr_init_avg["SiO4"][[1]][2])*(volume/(total_time)),
           ntr == "DOC" & Month == "May" ~ (conc- nutr_init_avg["DOC"][[1]][2])*(volume/(total_time)),
           ntr == "DON" & Month == "May" ~ (conc- nutr_init_avg["DON"][[1]][2])*(volume/(total_time)),
           ntr == "NOx" & Month == "May" ~ (conc- nutr_init_avg["NOx"][[1]][2])*(volume/(total_time)),
           TRUE ~ (conc- nutr_init_avg["TDN"][[1]][2])*(volume/(total_time))),
         dw = dw_leaves + dw_sponge + dw_rhizome2) #+ dw_rhizome
         
nutr_delta_ctrl <- nutr_rate |> 
    filter(Observation == "Control") |>
  group_by(light, Month, ntr) |> 
  summarise(avg_dnutr = mean(delta_cr),
            sd_dnutr = sd(delta_cr))

nutr_rate <- nutr_rate |> 
  filter(Observation %in% c("Posidonia", "Chondrilla", "Association")) |> 
  mutate(delta_cr_c = case_when(
           (ntr == "NH4") & (light == 0) & Month == "November" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "NH4" & nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "PO4" & (light == 0) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "PO4" & nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "NO2" & (light == 0) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "NO2"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "NO3" & (light == 0) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "NO3"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "SiO4" & (light == 0) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "SiO4"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "DOC" & (light == 0) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "DOC"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "DON" & (light == 0) & Month == "November" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "DON"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "NOx" & (light == 0) & Month == "November" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "NOx"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "TDN" & (light == 0) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "TDN"& nutr_delta_ctrl$Month == "November",4][[1]],
           (ntr == "NH4") & (light == 1) & Month == "November" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "NH4"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "PO4" & (light == 1) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "PO4"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "NO2" & (light == 1) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "NO2"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "NO3" & (light == 1) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "NO3"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "SiO4" & (light == 1) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "SiO4"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "DOC" & (light == 1) & Month == "November" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "DOC"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "DON" & (light == 1) & Month == "November" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "DON"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "NOx" & (light == 1) & Month == "November" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "NOx"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "TDN" & (light == 1) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "TDN"& nutr_delta_ctrl$Month == "November",4][[1]],
           (ntr == "NH4") & (light == 0) & Month == "May" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "NH4" & nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "PO4" & (light == 0) & Month == "May"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "PO4" & nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "NO2" & (light == 0) & Month == "May"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "NO2"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "NO3" & (light == 0) & Month == "May"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "NO3"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "SiO4" & (light == 0) & Month == "May"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "SiO4"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "DOC" & (light == 0) & Month == "May"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "DOC"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "DON" & (light == 0) & Month == "May" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "DON"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "NOx" & (light == 0) & Month == "May" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "NOx"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "TDN" & (light == 0) & Month == "May"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "TDN"& nutr_delta_ctrl$Month == "May",4][[1]],
           (ntr == "NH4") & (light == 1) & Month == "May" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "NH4"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "PO4" & (light == 1) & Month == "May"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "PO4"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "NO2" & (light == 1) & Month == "May"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "NO2"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "NO3" & (light == 1) & Month == "May"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "NO3"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "SiO4" & (light == 1) & Month == "May"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "SiO4"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "DOC" & (light == 1) & Month == "May" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "DOC"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "DON" & (light == 1) & Month == "May" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "DON"& nutr_delta_ctrl$Month == "May",4][[1]],
           TRUE ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "TDN"& nutr_delta_ctrl$Month == "May",4][[1]]),
         rate_ntr = (delta_cr_c)/(dw)
         )
```

### Dissolved Inorganic nutrients

```{r}
nutr_rate |>
   mutate(light = ifelse(light == 1, "Light", "Dark"),
          light = fct_relevel(light, "Light"),
          Month = fct_relevel(Month, "November"),
          Season = ifelse(Month == "November", "Autumn", "Spring"),
          Observation = fct_recode(Observation, 
                                  "Seagrass" = "Posidonia",
                                  "Sponge" = "Chondrilla"),
          Observation = fct_relevel(Observation,
                                   "Seagrass",
                                   "Sponge")
         ) |>
  filter(ntr %in% c("NOx", "PO4","NH4")) |>
  select(chamber, ntr, Season, light, rate_ntr, Observation) |> 
  ggplot(aes(x = Observation, y = rate_ntr, color = Season)) +
  geom_hline(yintercept = 0) +
  geom_point(size = 2.5,
             alpha = 0.5,
             position = position_dodge(width = 0.7)) +
  stat_summary(fun = mean,
               fun.min = function(x) mean(x) - sd(x),
               fun.max = function(x) mean(x) + sd(x),
               geom = "pointrange",
               #colour = "grey50",
               alpha = 0.5,
               shape = 3,
               position = position_dodge(width = 0.7)) +
  facet_grid(ntr~light, scales = "free") +
  #facet_wrap(light~ntr) +
  labs(y = expression(paste("Nutrient flux (",mu,"mol", " g"^"-1"," h"^"-1", ")"))) +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        strip.background = element_rect(fill = "white"),
        panel.grid.minor = element_blank(),
        text = element_text(size = 9),
        legend.position = "bottom")

ggsave("figures/fig4.svg",
       width = 169,
       height = 169,
       units = "mm")
```

```{r}
ntr_wide <-  nutr_rate |>
  select(chamber, Observation, Month, light, ntr, rate_ntr) |>
  pivot_wider(names_from = ntr,
              values_from = rate_ntr) 
  

mvabund_obj <- ntr_wide |> 
  select(nutr_names, -TDN, -NO3, -NO2, -SiO4) |> 
  mvabund()

meanvar.plot(mvabund_obj)

mod_nutr <- manylm(mvabund_obj ~ Observation*Month*light, 
                 data = ntr_wide)
plot(mod_nutr)
summary(mod_nutr)
anova(mod_nutr, p.uni="adjusted")
```

### Dissolved Organic nutrients

```{r}
nutr_rate |>
   mutate(light = ifelse(light == 1, "Light", "Dark"),
          light = fct_relevel(light, "Light"),
          Month = fct_relevel(Month, "November"),
          Season = ifelse(Month == "November", "Autumn", "Spring"),
          Observation = fct_recode(Observation, 
                                  "Seagrass" = "Posidonia",
                                  "Sponge" = "Chondrilla"),
          Observation = fct_relevel(Observation,
                                   "Seagrass",
                                   "Sponge")
         ) |>
  filter(ntr %in% c("DOC", "DON")) |>
  select(chamber, ntr, Season, light, rate_ntr, Observation) |> 
  ggplot(aes(x = Observation, y = rate_ntr, color = Season)) +
  geom_hline(yintercept = 0) +
  geom_point(size = 2.5,
             alpha = 0.5,
             position = position_dodge(width = 0.7)) +
  stat_summary(fun = mean,
               fun.min = function(x) mean(x) - sd(x),
               fun.max = function(x) mean(x) + sd(x),
               geom = "pointrange",
               #colour = "grey50",
               alpha = 0.5,
               shape = 3,
               position = position_dodge(width = 0.7)) +
  facet_grid(ntr~light, scales = "free") +
  labs(y = expression(paste("Nutrient flux (",mu,"mol", " g"^"-1"," h"^"-1", ")"))) +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        strip.background = element_rect(fill = "white"),
        panel.grid.minor = element_blank(),
        text = element_text(size = 9),
        legend.position = "bottom")

ggsave("figures/fig5.svg",
       width = 169,
       units = "mm")
```



