---
title: "3_nutrients"
format: html
editor_options: 
  chunk_output_type: console
---
# Preliminaries

```{r}
#| label: packages
#| message: false
#| warning: false
library(tidyverse)
library(vegan)
library(mvabund)
library(knitr)
library(patchwork)
library(pairwiseAdonis)
library(geosphere)
library(ggh4x)
#install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
library(DT)
```

```{r}
#| echo: false
#| label: directories
ifelse(!dir.exists("figures"),
       dir.create("figures"),
       FALSE)
```

```{r}
#| label: data
ponchos_raw <- read.csv("data/ponchos_master.csv")
```

In order to standardize the oxygen rates, we'll apply the following conversion factors:

$$ x \frac{mmol}{m^3} =  x \frac{\mu mol}{L}\cdot \text{Volume measured} (L) \cdot  \frac{1}{time (h)} \cdot \frac{1}{\text{biomass}(g)} \cdot = x\frac{\mu mol}{g_{biomass} \cdot h}$$

In this case, since the samples were collected from the entire volume, we can use 0.550 L as the common volume for all the samples

```{r}
#| echo: false
#| message: false
nutr_names <- c("NH4", "PO4", "NO2", "NO3", "NOx", "SiO4", "DOC", "TDN", "DON")

nutr_init_avg <- ponchos_raw |> 
  filter(chamber == "environment" & is.na(O2_0)) |>
  mutate(NOx = NO2 + NO3,
         DON = TDN - NOx - NH4) |> 
  group_by(Month) |> 
  summarise(across(all_of(nutr_names), mean))

nutr_init_sd <- ponchos_raw |> 
  filter(chamber == "environment" & is.na(O2_0)) |>
  mutate(NOx = NO2 + NO3,
         DON = TDN - NOx - NH4) |> 
   group_by(Month) |> 
  summarise(across(all_of(nutr_names), sd))
```

 Because some of the data include technical replicates, let's apply a custom
 function to select the replicates based on the combinations that minimize the coefficient of variance

```{r}
#| warning: false
source("R/trepf.R")
source("R/cv_get.R")
```

Then we can proceed to calculate the flux rates

```{r}
#| code-fold: true 
set.seed(1)

nutr_rate <- nutr_long |> 
  filter(!(Observation %in% c("t0", "rock"))) |>
  mutate(t0 = as.POSIXct(t0, format = "%H:%M"),
         tf = as.POSIXct(tf, format = "%H:%M"),
         total_time = as.numeric(tf - t0),
         Posidonia = as.factor(Posidonia),
         Chondrilla = as.factor(Chondrilla),
         light = as.factor(light),
         dw_rhizome2 = ifelse(
           !is.na(dw_rhizome),
           dw_rhizome,
           sample(ponchos_raw[ponchos_raw$dw_rhizome>0 & !is.na(ponchos_raw$dw_rhizome),"dw_rhizome"], 24)
         ),
         delta_cr = case_when(
           ntr == "NH4" & Month == "November" ~ (conc- nutr_init_avg["NH4"][[1]][1])*(volume/(total_time)),
           ntr == "PO4" & Month == "November" ~ (conc- nutr_init_avg["PO4"][[1]][1])*(volume/(total_time)),
           ntr == "NO2" & Month == "November" ~ (conc- nutr_init_avg["NO2"][[1]][1])*(volume/(total_time)),
           ntr == "NO3" & Month == "November" ~ (conc- nutr_init_avg["NO3"][[1]][1])*(volume/(total_time)),
           ntr == "SiO4" & Month == "November" ~ (conc- nutr_init_avg["SiO4"][[1]][1])*(volume/(total_time)),
           ntr == "DOC" & Month == "November" ~ (conc- nutr_init_avg["DOC"][[1]][1])*(volume/(total_time)),
           ntr == "DON" & Month == "November" ~ (conc- nutr_init_avg["DON"][[1]][1])*(volume/(total_time)),
           ntr == "NOx" & Month == "November" ~ (conc- nutr_init_avg["NOx"][[1]][1])*(volume/(total_time)),
           ntr == "TDN" & Month == "November" ~ (conc- nutr_init_avg["TDN"][[1]][1])*(volume/(total_time)),
           ntr == "NH4" & Month == "May" ~ (conc- nutr_init_avg["NH4"][[1]][2])*(volume/(total_time)),
           ntr == "PO4" & Month == "May" ~ (conc- nutr_init_avg["PO4"][[1]][2])*(volume/(total_time)),
           ntr == "NO2" & Month == "May" ~ (conc- nutr_init_avg["NO2"][[1]][2])*(volume/(total_time)),
           ntr == "NO3" & Month == "May" ~ (conc- nutr_init_avg["NO3"][[1]][2])*(volume/(total_time)),
           ntr == "SiO4" & Month == "May" ~ (conc- nutr_init_avg["SiO4"][[1]][2])*(volume/(total_time)),
           ntr == "DOC" & Month == "May" ~ (conc- nutr_init_avg["DOC"][[1]][2])*(volume/(total_time)),
           ntr == "DON" & Month == "May" ~ (conc- nutr_init_avg["DON"][[1]][2])*(volume/(total_time)),
           ntr == "NOx" & Month == "May" ~ (conc- nutr_init_avg["NOx"][[1]][2])*(volume/(total_time)),
           TRUE ~ (conc- nutr_init_avg["TDN"][[1]][2])*(volume/(total_time))),
         dw = dw_leaves + dw_sponge + dw_rhizome2) #+ dw_rhizome
         
nutr_delta_ctrl <- nutr_rate |> 
    filter(Observation == "Control") |>
  group_by(light, Month, ntr) |> 
  summarise(avg_dnutr = mean(delta_cr),
            sd_dnutr = sd(delta_cr))

nutr_rate <- nutr_rate |> 
  filter(Observation %in% c("Posidonia", "Chondrilla", "Association")) |> 
  mutate(delta_cr_c = case_when(
           (ntr == "NH4") & (light == 0) & Month == "November" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "NH4" & nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "PO4" & (light == 0) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "PO4" & nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "NO2" & (light == 0) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "NO2"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "NO3" & (light == 0) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "NO3"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "SiO4" & (light == 0) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "SiO4"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "DOC" & (light == 0) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "DOC"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "DON" & (light == 0) & Month == "November" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "DON"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "NOx" & (light == 0) & Month == "November" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "NOx"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "TDN" & (light == 0) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "TDN"& nutr_delta_ctrl$Month == "November",4][[1]],
           (ntr == "NH4") & (light == 1) & Month == "November" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "NH4"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "PO4" & (light == 1) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "PO4"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "NO2" & (light == 1) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "NO2"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "NO3" & (light == 1) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "NO3"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "SiO4" & (light == 1) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "SiO4"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "DOC" & (light == 1) & Month == "November" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "DOC"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "DON" & (light == 1) & Month == "November" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "DON"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "NOx" & (light == 1) & Month == "November" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "NOx"& nutr_delta_ctrl$Month == "November",4][[1]],
           ntr == "TDN" & (light == 1) & Month == "November"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "TDN"& nutr_delta_ctrl$Month == "November",4][[1]],
           (ntr == "NH4") & (light == 0) & Month == "May" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "NH4" & nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "PO4" & (light == 0) & Month == "May"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "PO4" & nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "NO2" & (light == 0) & Month == "May"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "NO2"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "NO3" & (light == 0) & Month == "May"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "NO3"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "SiO4" & (light == 0) & Month == "May"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "SiO4"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "DOC" & (light == 0) & Month == "May"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "DOC"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "DON" & (light == 0) & Month == "May" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "DON"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "NOx" & (light == 0) & Month == "May" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "NOx"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "TDN" & (light == 0) & Month == "May"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 0 & nutr_delta_ctrl$ntr == "TDN"& nutr_delta_ctrl$Month == "May",4][[1]],
           (ntr == "NH4") & (light == 1) & Month == "May" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "NH4"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "PO4" & (light == 1) & Month == "May"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "PO4"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "NO2" & (light == 1) & Month == "May"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "NO2"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "NO3" & (light == 1) & Month == "May"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "NO3"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "SiO4" & (light == 1) & Month == "May"~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "SiO4"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "DOC" & (light == 1) & Month == "May" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "DOC"& nutr_delta_ctrl$Month == "May",4][[1]],
           ntr == "DON" & (light == 1) & Month == "May" ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "DON"& nutr_delta_ctrl$Month == "May",4][[1]],
           TRUE ~ delta_cr- nutr_delta_ctrl[nutr_delta_ctrl$light == 1 & nutr_delta_ctrl$ntr == "TDN"& nutr_delta_ctrl$Month == "May",4][[1]]),
         rate_ntr = (delta_cr_c)/(dw)
         ) |> 
  mutate(rate_day = case_when(
    Month == "November" & light == 1 ~ rate_ntr*daylength(40, '2021-11-16'),
    Month == "November" & light == 0 ~ rate_ntr*(24-daylength(40, '2021-11-16')),
    Month == "May" & light == 1 ~ rate_ntr*daylength(40, '2022-05-16'),
    TRUE ~ rate_ntr*(24-daylength(40, '2022-05-16'))
    
  )
                             )
```

### Dissolved Organic nutrients
#### Mean and SD by combination of treatments

```{r}
#| echo: false
nutr_rate |>
   mutate(light = ifelse(light == 1, "Light", "Dark"),
          light = fct_relevel(light, "Light"),
          Month = fct_relevel(Month, "November"),
          Season = ifelse(Month == "November", "Autumn", "Spring"),
          Observation = fct_recode(Observation, 
                                  "Seagrass" = "Posidonia",
                                  "Sponge" = "Chondrilla"),
          Observation = fct_relevel(Observation,
                                   "Seagrass",
                                   "Sponge")
         ) |>
  filter(ntr %in% c("DOC", "DON")) |>
  select(chamber, ntr, Season, light, rate_ntr, Observation) |> 
  summarise(mean(rate_ntr),
              sd(rate_ntr),
            n(),
            .by = c(ntr, Season, Observation, light)
            ) |> 
  datatable(options = list(pageLength = 24))
```

#### Daily values by Season and Condition

```{r}
pooled_sd <- function(x, n) {sqrt(sum((x^2)/n))}

nutr_rate |>
  mutate(Treatment = ifelse(light == 1, "Light", "Dark"),
          Treatment = fct_relevel(Treatment, "Light"),
          Month = fct_relevel(Month, "November"),
          Season = ifelse(Month == "November", "Autumn", "Spring"),
          Observation = fct_recode(Observation, 
                                  "Seagrass" = "Posidonia",
                                  "Sponge" = "Chondrilla"),
          Observation = fct_relevel(Observation,
                                   "Seagrass",
                                   "Sponge")
         ) |> 
  filter(ntr %in% c("DOC", "DON")) |> 
  summarise(mean(rate_day), 
            sd(rate_day),
            n=n(),
            .by = c(Season, Treatment, Observation, ntr)) |> 
  summarise(sum_day = sum(`mean(rate_day)`),
            new_sd = pooled_sd(`mean(rate_day)`, n),
            .by = c(Season, Observation, ntr)
            ) |> 
  datatable()
```



```{r}
#| code-fold: true 
#| label: figure 4a

org_plot_A <- nutr_rate |>
   mutate(Treatment = ifelse(light == 1, "Light", "Dark"),
          Treatment = fct_relevel(Treatment, "Light"),
          Month = fct_relevel(Month, "November"),
          Season = ifelse(Month == "November", "Autumn", "Spring"),
          Observation = fct_recode(Observation, 
                                  "Seagrass" = "Posidonia",
                                  "Sponge" = "Chondrilla"),
          Observation = fct_relevel(Observation,
                                   "Seagrass",
                                   "Sponge")
         ) |>
  filter(ntr %in% c("DOC", "DON")) |>
  select(chamber, ntr, Season, Treatment, rate_ntr, Observation) |> 
  ggplot(aes(x = Observation, y = rate_ntr, color = Treatment)) +
  geom_hline(yintercept = 0) +
  geom_point(size = 2.5,
             alpha = 0.35,
             position = position_dodge(width = 0.7)) +
  stat_summary(fun = mean,
               fun.min = function(x) mean(x) - sd(x),
               fun.max = function(x) mean(x) + sd(x),
               geom = "pointrange",
               #colour = "grey50",
               alpha = 0.5,
               shape = 3,
               position = position_dodge(width = 0.7)) +
  facet_grid(ntr~Season, scales = "free") +
  scale_color_manual(values = c("blue", "black")) +
  labs(y = expression(paste("Nutrient flux (",mu,"mol", " g"^"-1"," h"^"-1", ")")),
       tag = "A"
       ) +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        strip.background = element_rect(fill = "white"),
        panel.grid.minor = element_blank(),
        text = element_text(size = 9),
        legend.position = "bottom")
```

```{r}
#| code-fold: true 
#| label: figure 4b
pooled_sd <- function(x, n) {sqrt(sum((x^2)/n))}

org_plot_B <- nutr_rate |>
  mutate(Treatment = ifelse(light == 1, "Light", "Dark"),
          Treatment = fct_relevel(Treatment, "Light"),
          Month = fct_relevel(Month, "November"),
          Season = ifelse(Month == "November", "Autumn", "Spring"),
          Observation = fct_recode(Observation, 
                                  "Seagrass" = "Posidonia",
                                  "Sponge" = "Chondrilla"),
          Observation = fct_relevel(Observation,
                                   "Seagrass",
                                   "Sponge")
         ) |> 
  filter(ntr %in% c("DOC", "DON")) |> 
  summarise(mean(rate_day), 
            sd(rate_day),
            n=n(),
            .by = c(Season, Treatment, Observation, ntr)) |> 
  summarise(sum_day = sum(`mean(rate_day)`),
            new_sd = pooled_sd(`mean(rate_day)`, n),
            .by = c(Season, Observation, ntr)
            ) |> 
  ggplot(aes(x = Observation, y = sum_day, color = Season)) +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(ymin = sum_day-new_sd,
                  ymax = sum_day+new_sd),
                  position = position_dodge(0.5),
                  shape = 3,
                  size = 1) +
  facet_grid(ntr~., scales = "free") +
  facetted_pos_scales(
    y = (ntr == "DON" ~ scale_y_continuous(limits = c(0, 40)))
  ) +
  labs(y = expression(paste("Nutrient flux (",mu,"mol", " g"^"-1"," day"^"-1", ")")),
       tag = "B"
       ) +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        strip.background = element_rect(fill = "white"),
        panel.grid.minor = element_blank(),
        text = element_text(size = 9),
        legend.position = "bottom")
```

```{r}
#| label: figure 4
org_plot_A + org_plot_B
#ggsave("figures/fig4.png", width = 169, height = 140, units = "mm")
```


### Dissolved Inorganic nutrients
#### Mean and SD by combination of treatments

```{r}
#| echo: false
#| fold: true
nutr_rate |>
   mutate(light = ifelse(light == 1, "Light", "Dark"),
          light = fct_relevel(light, "Light"),
          Month = fct_relevel(Month, "November"),
          Season = ifelse(Month == "November", "Autumn", "Spring"),
          Observation = fct_recode(Observation, 
                                  "Seagrass" = "Posidonia",
                                  "Sponge" = "Chondrilla"),
          Observation = fct_relevel(Observation,
                                   "Seagrass",
                                   "Sponge")
         ) |>
  filter(ntr %in% c("NOx", "PO4","NH4")) |>
  select(chamber, ntr, Season, light, rate_ntr, Observation) |> 
  summarise(mean(rate_ntr),
              sd(rate_ntr),
            n(),
            .by = c(ntr, Season, Observation, light)
            ) |> 
  datatable(options = list(pageLength = 36))
```

#### Daily values by Season and Condition

```{r}
#| fold: true
nutr_rate |>
  mutate(Treatment = ifelse(light == 1, "Light", "Dark"),
          Treatment = fct_relevel(Treatment, "Light"),
          Month = fct_relevel(Month, "November"),
          Season = ifelse(Month == "November", "Autumn", "Spring"),
          Observation = fct_recode(Observation, 
                                  "Seagrass" = "Posidonia",
                                  "Sponge" = "Chondrilla"),
          Observation = fct_relevel(Observation,
                                   "Seagrass",
                                   "Sponge")
         ) |> 
  filter(ntr %in% c("NOx", "PO4","NH4")) |>
  summarise(mean(rate_day), 
            sd(rate_day),
            n=n(),
            .by = c(Season, Treatment, Observation, ntr)) |> 
  summarise(sum_day = sum(`mean(rate_day)`),
            new_sd = pooled_sd(`mean(rate_day)`, n),
            .by = c(Season, Observation, ntr)
            ) |>
  datatable()
```


```{r}
#| code-fold: true 
#| label: figure-5a
inorg_plot_A <- nutr_rate |>
   mutate(Treatment = ifelse(light == 1, "Light", "Dark"),
          Treatment = fct_relevel(Treatment, "Light"),
          Month = fct_relevel(Month, "November"),
          Season = ifelse(Month == "November", "Autumn", "Spring"),
          Observation = fct_recode(Observation, 
                                  "Seagrass" = "Posidonia",
                                  "Sponge" = "Chondrilla"),
          Observation = fct_relevel(Observation,
                                   "Seagrass",
                                   "Sponge")
         ) |>
  filter(ntr %in% c("NOx", "PO4","NH4")) |>
  select(chamber, ntr, Season, Treatment, rate_ntr, Observation) |> 
  ggplot(aes(x = Observation, y = rate_ntr, color = Treatment)) +
  geom_hline(yintercept = 0) +
  geom_point(size = 2.5,
             alpha = 0.5,
             position = position_dodge(width = 0.7)) +
  stat_summary(fun = mean,
               fun.min = function(x) mean(x) - sd(x),
               fun.max = function(x) mean(x) + sd(x),
               geom = "pointrange",
               #colour = "grey50",
               alpha = 0.5,
               shape = 3,
               position = position_dodge(width = 0.7)) +
  facet_grid(ntr~Season, scales = "free") +
  scale_color_manual(values = c("blue", "black")) +
  labs(y = expression(paste("Nutrient flux (",mu,"mol", " g"^"-1"," h"^"-1", ")")),
       tag = "A"
       ) +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        strip.background = element_rect(fill = "white"),
        panel.grid.minor = element_blank(),
        text = element_text(size = 9),
        legend.position = "bottom")
```


```{r}
#| code-fold: true 
#| label: figure-5b

inorg_plot_B <- nutr_rate |>
  mutate(Treatment = ifelse(light == 1, "Light", "Dark"),
          Treatment = fct_relevel(Treatment, "Light"),
          Month = fct_relevel(Month, "November"),
          Season = ifelse(Month == "November", "Autumn", "Spring"),
          Observation = fct_recode(Observation, 
                                  "Seagrass" = "Posidonia",
                                  "Sponge" = "Chondrilla"),
          Observation = fct_relevel(Observation,
                                   "Seagrass",
                                   "Sponge")
         ) |> 
  filter(ntr %in% c("NOx", "PO4","NH4")) |>
  summarise(mean(rate_day), 
            sd(rate_day),
            n=n(),
            .by = c(Season, Treatment, Observation, ntr)) |> 
  summarise(sum_day = sum(`mean(rate_day)`),
            sum_sd_day = sum(`sd(rate_day)`),
            new_sd = pooled_sd(`mean(rate_day)`, n),
            .by = c(Season, Observation, ntr)
            ) |> 
  ggplot(aes(x = Observation, y = sum_day, color = Season)) +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(ymin = sum_day-new_sd,
                  ymax = sum_day+new_sd),
                  position = position_dodge(0.5),
                  shape = 3,
                  size = 1) +
  facet_grid(ntr~., scales = "free") +
  facetted_pos_scales(
    y = list(ntr == "NH4" ~ scale_y_continuous(limits = c(-5, 5)),
             ntr == "PO4" ~ scale_y_continuous(limits = c(-0.1, 1)))
  ) +
  labs(y = expression(paste("Nutrient flux (",mu,"mol", " g"^"-1"," day"^"-1", ")")),
       tag = "B"
       ) +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        strip.background = element_rect(fill = "white"),
        panel.grid.minor = element_blank(),
        text = element_text(size = 9),
        legend.position = "bottom")

```

```{r}
inorg_plot_A + inorg_plot_B
#ggsave("figures/fig5.svg", width = 169, height = 169, units = "mm")
```


## Statistical analyses

```{r}
#| label: Nutrient individual permanova

ntr_wide <-  nutr_rate |>
  select(chamber, Observation, Month, light, ntr, rate_ntr) |>
  pivot_wider(names_from = ntr,
              values_from = rate_ntr) 

ntr_select <- ntr_wide |> 
  select(nutr_names, -TDN, -NO3, -NO2, -SiO4) |> 
  colnames()

ntr_perm <- lapply(ntr_select, function(x){
  result <- adonis2(formula = ntr_wide[,x] ~ Observation * Month * light,
                    data = ntr_wide,
                    method = "euclidean"
                    )
})

names(ntr_perm) <- ntr_select
```

### supp table 3

```{r}
#| label: supplementary table 3


ntr_summ <- ntr_wide |> 
  mutate(
    Condition = ifelse(ntr_wide$light == 1, "Light", "Dark"),
    Season = ifelse(ntr_wide$Month == "November", "Autumn", "Spring"),
    inter = paste(Season,"-",Condition, sep = "")
    ) |> 
  select(Observation, Season, Condition, ntr_select) |> 
  #group_by(Observation, Season, Condition) |> 
  summarize(
    across(is.numeric, list(mean = mean, sd = sd, 'n' = ~n())),
    .by = c(Observation, Season, Condition)
  ) 

ntr_summ |> 
  datatable()


sink("supplementary/supp_table_3.md")
ntr_summ |> 
    kable(format = "markdown",
        caption = "Organic and inorganic nutrient mean, sd and n")

sink()
```

### Pairwise-tests

```{r}
#| label: DOC

ntr_wide |> 
  mutate(
    Treatment = ifelse(ntr_wide$light == 1, "Light", "Dark"),
    Season = ifelse(ntr_wide$Month == "November", "Autumn", "Spring"),
    inter = paste(Season,"-",Treatment, sep = "")
    ) |> 
  select(DOC, inter) |> 
  {\(x) aov(DOC~inter,
                         data = x)}() |> 
  TukeyHSD() 
```

```{r}
#| label: NH4
ntr_wide |> 
  mutate(
    Treatment = ifelse(ntr_wide$light == 1, "Light", "Dark"),
    Season = ifelse(ntr_wide$Month == "November", "Autumn", "Spring"),
    inter = paste(Season,"-",Treatment, sep = "")
    ) |> 
  select(NH4, inter) |> 
  {\(x) aov(NH4~inter,
                         data = x)}() |> 
  TukeyHSD() 
```

```{r}
#| label: PO4
ntr_wide |> 
  select(PO4, Observation) |> 
  {\(x) aov(PO4~Observation,
                         data = x)}() |> 
  TukeyHSD() 
```

```{r}
#| label: NOx
ntr_wide |> 
  mutate(
    Treatment = ifelse(ntr_wide$light == 1, "Light", "Dark"),
    Season = ifelse(ntr_wide$Month == "November", "Autumn", "Spring"),
    inter = paste(Season,"-",Treatment, sep = "")
    ) |> 
  select(NOx, inter) |> 
  {\(x) aov(NOx~inter,
                         data = x)}() |> 
  TukeyHSD() 

ntr_wide |> 
  mutate(
    Treatment = ifelse(ntr_wide$light == 1, "Light", "Dark"),
    inter = paste(Observation,"-",Treatment, sep = "")
    ) |> 
  select(NOx, inter) |> 
  {\(x) aov(NOx~inter,
                         data = x)}() |> 
  TukeyHSD()
```

